# Sample sizes of Trondheim dataset

* 70.082 individual buildings.
* 47.853 cadastral plots.
* 39.049 cadastral plots containing buildings.
  * Q1/Q2/Q3: 375/674/1121 m^2
  * Mean: 6862 m^2

# Cadastre bounding box

## Distribution of pixels size in cadastre bounding box

The bounding box is forced to have no angular orientation
with respect to the north-bound direction.

```{r}
library(tidyverse)
bbox_stats <- read_delim("../data/bbox_stats.csv", delim=",")
bbox_stats %>%
  filter(area < quantile(area, 0.95)) %>%
  ggplot() +
    aes(x = sqrt(area * 16)) +
    geom_histogram(binwidth=4) +
    geom_vline(
      aes(
        xintercept = median(sqrt(area * 16))
      )
    ) +
    xlab("sqrt(total pixels)")
```

## Distribution of width and height of cadastre bounding box in pixels

```{r}
bbox_stats %>%
  gather(key, value, width, height) %>%
  filter(value < quantile(value, 0.95)) %>%
  ggplot() +
    aes(x = value * 4) +
    geom_histogram(binwidth=1) +
    geom_vline(
      aes(
        xintercept = median(value * 4)
      )
    ) +
    facet_wrap(~key) +
    xlab("Pixels")
```

# Cadastre minimal rectangle

The minimal rectangle can have arbitrary orientations.

## Distribution of pixels size in cadastre minimal oriented rectangle

```{r}
library(tidyverse)
minrec_stats <- read_delim("../data/minrec_stats.csv", delim=",")
minrec_stats %>%
  filter(area < quantile(area, 0.95)) %>%
  ggplot() +
    aes(x = sqrt(area * 16)) +
    geom_histogram(binwidth=4) +
    geom_vline(
      aes(
        xintercept = median(sqrt(area * 16))
      )
    ) +
    xlab("sqrt(total pixels)")
```

## Distribution of width and height of cadastre minimal rectangle

```{r}
minrec_stats %>%
  gather(key, value, width, height) %>%
  filter(value < quantile(value, 0.95)) %>%
  ggplot() +
    aes(x = value * 4) +
    geom_histogram(binwidth=1) +
    geom_vline(
      aes(
        xintercept = median(value * 4)
      )
    ) +
    facet_wrap(~key) +
    xlab("Pixels")
```

## Distribution of orientation

```{r}
minrec_stats %>%
  ggplot() +
    aes(x = angle * pi / 180, y = ..density..) +
    geom_histogram(binwidth=1) +
    coord_polar(start = pi) +
    theme(
      aspect.ratio = 1
    )
```

